---
title: "Project 1"
author: "Lidya Tadesse"
date: "9/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Explore the Data
1. How many songs are in each genre? <br/>

|                 edm = 6043 
|                 latin = 5155 
|                 pop = 5507 
|                 r&b = 5431 
|                 rap = 5746 
|                 rock = 4951 

<br/>

Code:

```{r setup2, include=TRUE}
require(tidytuesdayR)
require(tidyverse)
tuesdata <- tidytuesdayR::tt_load('2020-01-21')
spotify_songs <- tuesdata$spotify_songs 
```

``` {r}
spotify_songs %>% count(playlist_genre)

```

<br/><br/>

2. What is average value of energy and acousticness in the latin genre in this dataset? <br/>

|                 mean energy = 0.708
|                 mean acousticness = 0.211 


<br/>

Code:

``` {r}
genre <- group_by (spotify_songs, playlist_genre)
summarise(genre, avg_energy = mean(energy), avg_acoustic = mean(acousticness))
```

<br/><br/>

3. Calculate the average duration of song (in minutes) across all subgenres. Which subgenre has the longest song on average? <br/>


|                 New Jack Swing is the subgenre with the longest songs. On average, they are 275128.6 ms in length.

<br/>

Code:

``` {r}
subgenre <- group_by(spotify_songs, playlist_subgenre)
subgenre_durations <- summarize(subgenre, duration_ms = mean(duration_ms, na.rm = TRUE) )
```

<br/><br/>

4. Make two boxplots side-by-side of the danceability of songs stratifying by whether a song has a fast or slow tempo. Define fast tempo as any song that has a tempo above its median value. On average, which songs are more danceable?


|               Median tempo is 121.98. 

<br/><br/>

``` {r} 
danceability <- spotify_songs %>% mutate(tempocat= case_when(tempo > 121.98 ~ "high", TRUE ~ "low"))
```

``` {r danceability_boxplot, eval = TRUE, echo = TRUE}
require(ggplot2)
```

```
qplot(tempocat, danceability, data = danceability, geom="boxplot", fill=tempocat, main="Danceability by Tempo Category", xlab="Tempo", ylab="Danceability")
```       

<br/>

|       On average, lower tempo songs are more danceable than higher tempo songs.

<br/><br/><br/>

## Part 2: Convert nontidy data into tidy data

1. Select only unique distinct rows from the dataset based on the track_name and track_artist columns.

``` {r}
unique <- distinct(spotify_songs, track_name, track_artist, .keep_all = FALSE)
```

<br/>

2. Add a new column called year_released listing just the year that the song was released.

``` {r}
spotify_songs2 <- spotify_songs %>% separate(col="track_album_release_date", into=c("year", "month", "date"), sep="-")
spotify_songs2 <- distinct(spotify_songs2, track_name, track_artist, year, .keep_all = FALSE)
unique_year <- left_join(x = unique, y = spotify_songs2, by = c("track_name", "track_artist"))
colnames(unique_year)[colnames(unique_year) == 'year'] <- 'year_released'
```

3. Keep only songs that were released in or after 1980.

``` {r}
post_1980 <- filter(unique_year, year_released >= 1980)
```

<br/>

4. Add a new column with the duration of the song in minutes

``` {r}

duration <- spotify_songs %>% 
  select(track_name, track_artist, duration_ms)
duration <- distinct(duration, track_name, track_artist, duration_ms, .keep_all = FALSE)
update_v4 <- left_join(x = post_1980, y = duration, by = c("track_name", "track_artist"))
```
<br/>

5. For each year_released, calculate the mean of at least 6 of the audio features (e.g. danceability, energy, loudness, etc), or descriptors (e.g. tempo, duration in minutes, etc).

``` {r}

#Adding the audio feature variables

#energy
energy <- spotify_songs %>% 
  select(track_name, track_artist, energy)
energy <- distinct(energy, track_name, track_artist, energy, .keep_all = FALSE)
update_v5 <- left_join(x = update_v4, y = energy, by = c("track_name", "track_artist"))

#key
key <- spotify_songs %>% 
  select(track_name, track_artist, key)
key <- distinct(key, track_name, track_artist, key, .keep_all = FALSE)
update_v6 <- left_join(x = update_v5, y = key, by = c("track_name", "track_artist"))

#acousticness
acoustic <- spotify_songs %>% 
  select(track_name, track_artist, acousticness)
acoustic <- distinct(acoustic, track_name, track_artist, acousticness, .keep_all = FALSE)
update_v7 <- left_join(x = update_v6, y = acoustic, by = c("track_name", "track_artist"))

#liveness
live <- spotify_songs %>% 
  select(track_name, track_artist, liveness)
live <- distinct(live, track_name, track_artist, liveness, .keep_all = FALSE)
update_v8 <- left_join(x = update_v7, y = live, by = c("track_name", "track_artist"))

#tempo
tempo <- spotify_songs %>% 
  select(track_name, track_artist, tempo)
tempo <- distinct(tempo, track_name, track_artist, tempo, .keep_all = FALSE)
update_v9 <- left_join(x = update_v8, y = tempo, by = c("track_name", "track_artist"))

#speechiness
speech <- spotify_songs %>% 
  select(track_name, track_artist, speechiness)
speech <- distinct(speech, track_name, track_artist, speechiness, .keep_all = FALSE)
update_v10 <- left_join(x = update_v9, y = speech, by = c("track_name", "track_artist"))

```
<br/>
``` {r}

#Creating table

group_year <- group_by(update_v10, year_released)
summarise(group_year, avg_energy = mean(energy), avg_key = mean(key), avg_acoustic = mean(acousticness), avg_live = mean(liveness), avg_dur = mean(duration_ms), avg_tempo = mean(tempo), avg_speech = mean(speechiness))

```
