---
title: "Project 1"
author: "Lidya Tadesse"
date: "9/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Explore the Data
1. How many songs are in each genre? <br/>

|                 edm = 6043 
|                 latin = 5155 
|                 pop = 5507 
|                 r&b = 5431 
|                 rap = 5746 
|                 rock = 4951 

<br/>

2. What is average value of energy and acousticness in the latin genre in this dataset? <br/>

|                 mean energy = 0.7083125
|                 mean acousticness = 0.2109204  


<br/>

3. Calculate the average duration of song (in minutes) across all subgenres. Which subgenre has the longest song on average? <br/>


|                 New Jack Swing is the subgenre with the longest songs. On average, they are 275128.6 ms in length.

<br/>

4. Make two boxplots side-by-side of the danceability of songs stratifying by whether a song has a fast or slow tempo. Define fast tempo as any song that has a tempo above its median value. On average, which songs are more danceable?


|               Median tempo is 121.98. 

<br/><br/>

``` {r moresetup, include = FALSE} 
require(tidytuesdayR)
require(tidyverse)
tuesdata <- tidytuesdayR::tt_load('2020-01-21')
spotify_songs <- tuesdata$spotify_songs 
danceability <- spotify_songs %>% mutate(tempocat= case_when(tempo < 121.98 ~ "1", TRUE ~ "0"))
require(ggplot2)
```

``` {r danceability_boxplot, eval = TRUE, echo = FALSE}
qplot(tempocat, danceability, data = danceability, geom="boxplot", fill=tempocat, main="Danceability by Tempo Category", xlab="Tempo", ylab="Danceability")
```       













