---
title: "Project 1"
author: "Lidya Tadesse"
date: "9/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part 1: Explore the Data
1. How many songs are in each genre? <br/>

|                 edm = 6043 
|                 latin = 5155 
|                 pop = 5507 
|                 r&b = 5431 
|                 rap = 5746 
|                 rock = 4951 

<br/>

Code:

```{r setup2, include=TRUE}
require(tidytuesdayR)
require(tidyverse)
tuesdata <- tidytuesdayR::tt_load('2020-01-21')
spotify_songs <- tuesdata$spotify_songs 
```

``` {r}
spotify_songs %>% count(playlist_genre)

```

<br/><br/>

2. What is average value of energy and acousticness in the latin genre in this dataset? <br/>

|                 mean energy = 0.708
|                 mean acousticness = 0.211 


<br/>

Code:

``` {r}
genre <- group_by (spotify_songs, playlist_genre)
summarise(genre, avg_energy = mean(energy), avg_acoustic = mean(acousticness))
```

<br/><br/>

3. Calculate the average duration of song (in minutes) across all subgenres. Which subgenre has the longest song on average? <br/>


|                 New Jack Swing is the subgenre with the longest songs. On average, they are 275128.6 ms in length.

<br/>

Code:

``` {r}
subgenre <- group_by(spotify_songs, playlist_subgenre)
subgenre_durations <- summarize(subgenre, duration_ms = mean(duration_ms, na.rm = TRUE) )
```

<br/><br/>

4. Make two boxplots side-by-side of the danceability of songs stratifying by whether a song has a fast or slow tempo. Define fast tempo as any song that has a tempo above its median value. On average, which songs are more danceable?


|               Median tempo is 121.98. 

<br/><br/>

``` {r} 
danceability <- spotify_songs %>% mutate(tempocat= case_when(tempo > 121.98 ~ "high", TRUE ~ "low"))
```

``` {r danceability_boxplot, eval = TRUE, echo = TRUE}
require(ggplot2)
```

```
qplot(tempocat, danceability, data = danceability, geom="boxplot", fill=tempocat, main="Danceability by Tempo Category", xlab="Tempo", ylab="Danceability")
```       

<br/>

|       On average, lower tempo songs are more danceable than higher tempo songs.

<br/><br/><br/>

## Part 2: Convert nontidy data into tidy data

1. Select only unique distinct rows from the dataset based on the track_name and track_artist columns.

``` {r}
unique <- distinct(spotify_songs, spotify_songs$track_name, spotify_songs$track_artist, .keep_all = FALSE)
```

<br/>

2. Add a new column called year_released listing just the year that the song was released.






